### 2016-06-28 solar_letters_cover.tex

\documentclass[12pt]{article}
\usepackage{setspace}
\usepackage[margin=1in]{geometry}
\usepackage{mathtools}
\usepackage{natbib} %for citet and citep
\usepackage{syntonly}
\usepackage{esdiff} %for writing partial derivatives
\usepackage{url} %for inserting urls
\usepackage{placeins}
\usepackage{textcomp} 
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}

%\doublespacing % from package setspacs

\title{Lease or Buy? Quality differences and asymmetric information in California solar panels}

\date{\today}

\author{Johannes Mauritzen \\ BI Norwegian Business School \\ Trondheim Campus \\ Department of Accounting, Auditing and Business Analytics \\ johannes.mauritzen@bi.no\\\url{http://jmaurit.github.io}}

\begin{document}
 \begin{spacing}{1} %sets spacing to single for title page
	\maketitle

\begin{abstract}
 I provide one of few empirical case studies of asymmetric information of quality in a durable good. This study is unique in observing quality directly. I use data from approximately 1000 rooftop solar photovoltaic systems in California with 4 to 5 years of monthly production data. Using a Bayesian hierarchical model, I find evidence that leased systems show a higher level of quality than those bought out-right.

 JEL Codes: L15, Q42
\end{abstract}

% \thanks{*I would like to thank...}
% 
 \end{spacing}
 \end{document}