### 2019-03-29 ejor_submission/R&R/replyToEditor.txt
update RR

Reply to editor

As you can see from the reports enclosed, the referees hold conflicting views on your paper. In these circumstances, it has been decided that your paper should be reconsidered for publication in the EUROPEAN JOURNAL OF OPERATIONAL RESEARCH if you will revise it to meet the enclosed referees' comments.

The most critical comments made me doubt whether I should reconsider your paper, so I need convincing arguments from you and a thorough revision to counterbalance the critical comments.

When resubmitting the revised paper, please provide your answers to the referees, separately for each referee.  Please also ensure that you have cited recent and relevant publications in EJOR and other OR journals.

I am most grateful for your interest and hope that you will not be deterred from resubmitting this paper in the very near future.

You are advised to revise and resubmit your manuscript by the Sep 19, 2019. After this time the review process of this submission will be terminated, by withdrawing your paper from the Elsevier Editorial System, and the revised paper will have to be submitted as a new submission.